The Chemistry Development Kit (CDK) \cdkversion{}: atom typing, depiction, molecular formulas, and substructure searching

### 2017-04-29 article.tex
Fixed a few typos

OSGi bundles are available for the CDK too, which are used by e.g.
Bioclipse~\cite{spjuth2007bioclipse,spjuth2009bioclipse} and
KNIME~\cite{Beisken2013}. However, because CDK Java packages are occassionally
split between CDK modules, the CDK currently needs to be bundled as a single
OSGi jar. The bundle is available from \url{http://pele.farmbio.uu.se/bioclipse/cdk/cdk-1.5.13/}.
This Java package and bundle incompatibilities are currently being explored and
constitutes an area where improvements can be done on modularization.

@@

\textbf{-type maccs}: The CDK MACCS fingerprint uses 166 keys to encode
features of a structure and can be used for similarity searching. This encoding
uses different aspects of the library including ring perception and the new
aromaticity perception but the speed is primarily dependant on SMARTS matching 
performance. Generating the fingerprint previously took almost 1 day for ChEMBL 
and 2 hours for ChEBI. This has been reduced to less than 
15 minutes for ChEMBL and less than 30 seconds for ChEBI.

@@

\begin{figure}[h!]
  \caption{\csentence{Relative storage of stereochemistry, the \textbf{type} and
    \textbf{focus} of stereochemsitry are fixed for a given
    stereocenter description but the \textbf{carriers} and
    \textbf{configuration} are relative.} 
  The multiple rows for each
  stereochemistry type are different internal representation that
  would be considered equivalent. In the tetrahedral types,
  hydrogens may be suppressed in a molecular graph so the
  \textbf{focus} is reused in the \textbf{carriers} list as a
  placeholder.}
\centering
%\includegraphics{img/stereodesc_annotated.pdf}
\label{fig:stereodatastructure}
\end{figure}

@@

    \begin{minipage}{1\textwidth}
    \renewcommand*{\thempfootnote}{\fnsymbol{mpfootnote}}
    \centering
    \begin{tabular}{lp{3cm}p{3cm}l}
  \textbf{Module}            & \textbf{Description}  & \textbf{Major Changes} & \textbf{Dependencies} \\ \hline
  interfaces                 & Interfaces for the data models. & & Vecmath 1.5.2 \\ \hline
  core                       & Core functionality.             & & Google Guava 17.0 \\ \hline % FIXME: how to add https://github.com/google/guava ?
  standard                   & Common functionality.           & & \\ \hline
  render                     & Graphical rendering.            & Redesigned to make it more modular and support multiple widget toolkits, like AWT and SWT. & \\ \hline
  isomorphism                & Isomorphism and substructure searching. & & \\ \hline
  atomtype                   & Various non-core atom type schemes.     & Unified approach where atom typing is separated from other algorithms. & \\ \hline
  ioformats                  & Definitions of (chemical) input/output formats. & & \\ \hline
  io                         & Readers and writers for input/output formats.  & The molfile reader has been rewritten and supports in atom types defind in the specification. & XPP3 1.1.4c \\ \hline
  iordf                      & Stores data models as in the Resource Description Framework serialization formats. & New. & Jena 2.7.4 \\ \hline
  inchi                      & IUPAC International Chemical Identifier support. & & JNI-InChI 0.8~\cite{Spjuth2013}  \\ \hline
  libiocml                   & Writer for the Chemical Markup Language format. & & XOM 1.2.5, CMLXOM 3.1~\cite{Murray-Rust2011} \\ \hline
  sdg                        & Structure diagram generation.  & Much improved overlap resolution. & \\ \hline
  smiles                     & Reading and writing in the SMILES format. & SMILES support performance and coverage is greatly improved. & Beam 0.9.1~\cite{Beam} \\ \hline
  smarts                     & Substructure searching with the SMARTS format. & & Beam 0.9.1~\cite{Beam} \\ \hline
  hash                       & Molecular hash codes~\cite{Ihlenfeldt94}. & & \\ \hline
  formula                    & Chemical formula support. & New. & \\ \hline
  fingerprint                & Calculate fingerprints. & Many new fingerprint types (see text). & Apache Commons Math 3.1.1 \\ \hline
  qsar and qsarmolecular                       & Molecular descriptors.  & & XOM 1.2.5, JAMA 1.0.3~\cite{Hicklin2012} \\ \hline
  signatures                 & Calculation of molecular and atomic signatures. & & Signatures 1.1 \\ \hline
    \end{tabular}
    \end{minipage}

>>>>

OSGi bundles are available for the CDK too, which are used by e.g.
Bioclipse~\cite{spjuth2007bioclipse,spjuth2009bioclipse} and
KNIME~\cite{Beisken2013}. However, because CDK Java packages are occasionally
split between CDK modules, the CDK currently needs to be bundled as a single
OSGi jar. The bundle is available from \url{http://pele.farmbio.uu.se/bioclipse/cdk/cdk-1.5.13/}.
This Java package and bundle incompatibilities are currently being explored and
constitutes an area where improvements can be done on modularization.

@@

\textbf{-type maccs}: The CDK MACCS fingerprint uses 166 keys to encode
features of a structure and can be used for similarity searching. This encoding
uses different aspects of the library including ring perception and the new
aromaticity perception but the speed is primarily dependent on SMARTS matching 
performance. Generating the fingerprint previously took almost 1 day for ChEMBL 
and 2 hours for ChEBI. This has been reduced to less than 
15 minutes for ChEMBL and less than 30 seconds for ChEBI.

@@

\begin{figure}[h!]
  \caption{\csentence{Relative storage of stereochemistry, the \textbf{type} and
    \textbf{focus} of stereochemistry are fixed for a given
    stereocenter description but the \textbf{carriers} and
    \textbf{configuration} are relative.} 
  The multiple rows for each
  stereochemistry type are different internal representation that
  would be considered equivalent. In the tetrahedral types,
  hydrogens may be suppressed in a molecular graph so the
  \textbf{focus} is reused in the \textbf{carriers} list as a
  placeholder.}
\centering
%\includegraphics{img/stereodesc_annotated.pdf}
\label{fig:stereodatastructure}
\end{figure}

@@

    \begin{minipage}{1\textwidth}
    \renewcommand*{\thempfootnote}{\fnsymbol{mpfootnote}}
    \centering
    \begin{tabular}{lp{3cm}p{3cm}l}
  \textbf{Module}            & \textbf{Description}  & \textbf{Major Changes} & \textbf{Dependencies} \\ \hline
  interfaces                 & Interfaces for the data models. & & Vecmath 1.5.2 \\ \hline
  core                       & Core functionality.             & & Google Guava 17.0 \\ \hline % FIXME: how to add https://github.com/google/guava ?
  standard                   & Common functionality.           & & \\ \hline
  render                     & Graphical rendering.            & Redesigned to make it more modular and support multiple widget toolkits, like AWT and SWT. & \\ \hline
  isomorphism                & Isomorphism and substructure searching. & & \\ \hline
  atomtype                   & Various non-core atom type schemes.     & Unified approach where atom typing is separated from other algorithms. & \\ \hline
  ioformats                  & Definitions of (chemical) input/output formats. & & \\ \hline
  io                         & Readers and writers for input/output formats.  & The molfile reader has been rewritten and supports atom types defined in the specification. & XPP3 1.1.4c \\ \hline
  iordf                      & Stores data models as in the Resource Description Framework serialization formats. & New. & Jena 2.7.4 \\ \hline
  inchi                      & IUPAC International Chemical Identifier support. & & JNI-InChI 0.8~\cite{Spjuth2013}  \\ \hline
  libiocml                   & Writer for the Chemical Markup Language format. & & XOM 1.2.5, CMLXOM 3.1~\cite{Murray-Rust2011} \\ \hline
  sdg                        & Structure diagram generation.  & Much improved overlap resolution. & \\ \hline
  smiles                     & Reading and writing in the SMILES format. & SMILES support performance and coverage is greatly improved. & Beam 0.9.1~\cite{Beam} \\ \hline
  smarts                     & Substructure searching with the SMARTS format. & & Beam 0.9.1~\cite{Beam} \\ \hline
  hash                       & Molecular hash codes~\cite{Ihlenfeldt94}. & & \\ \hline
  formula                    & Chemical formula support. & New. & \\ \hline
  fingerprint                & Calculate fingerprints. & Many new fingerprint types (see text). & Apache Commons Math 3.1.1 \\ \hline
  qsar and qsarmolecular                       & Molecular descriptors.  & & XOM 1.2.5, JAMA 1.0.3~\cite{Hicklin2012} \\ \hline
  signatures                 & Calculation of molecular and atomic signatures. & & Signatures 1.1 \\ \hline
    \end{tabular}
    \end{minipage}
